Poker Chips, Earthquakes, and Computation

### 2014-09-26 draft0.tex
rearranged section order
heavily revised the introduction

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage[margin=0.9in]{geometry}
%\usepackage{amsfonts}
%\usepackage{amsthm}
\usepackage{subfigure}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{cite}
\usepackage{hyperref}
\usepackage{url}
\setlength{\parindent}{0cm}
\urlstyle{same}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

@@

In this paper, we investigate the following problem:

@@

We also consider extensions of this problem set up where the earthquake hits all squares on the board at once. We refer to the former type of earthquake as $\textbf{local earthquakes}$ and earthquakes that hit the entire board as $\textbf{global earthquakes}$. \\

The paper is organized as follows. In Section~\ref{Definitions and Notation}, we formalize the terminology and operations we use to analyze the game. In Section~\ref{Stability of Finite Boards}, we introduce the idea of a potential function defined on boards to prove that all boards with a finite number of starting chips eventually stabilize. We then modify the potential function to work for infinite boards and we explore its consequences in Section~\ref{Acyclicity of Infinite Boards}. In Section~\ref{Order of Redistribution}, we show that the order in which local earthquakes happens does not matter, thus answering a question in the original problem statement. We then shift our approach in Section~\ref{Time until Stability} and present results from our simulations of repeated earthquakes. In particular, we look at the starting arrangement where $n$ chips all start on the same square, and we conjecture based on experimental evidence that such arragnements stabilize after $\Theta(n \log n)$ global earthquakes. In Section~\ref{Board Variants: n-Trees}, we consider what happens when alternative board setups are considered. Finally, in Section~\ref{Computing with Chips} we present a model of computation based on the problem set up, and show how to construct simple Boolean circuits and prove some impossibility results. 
\section{Definitions and Notation}
\label{Definitions and Notation}

>>>>

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage[margin=0.9in]{geometry}
%\usepackage{amsfonts}
%\usepackage{amsthm}
\usepackage{subfigure}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{cite}
\usepackage{hyperref}
\usepackage{url}
\urlstyle{same}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

@@

\noindent In this paper, we investigate the following problem:

@@

We also consider extensions of this problem where the earthquake hits all squares on the board at once. We will refer to earthquakes that hit only one square as $\emph{local earthquakes}$ and earthquakes that hit the entire board as $\emph{global earthquakes}$.

The paper is organized so that purely mathematical results addressing the original problem statement come first. In Section~\ref{Definitions and Notation}, we formalize the terminology and operations we use to analyze the problem. In particular, we formalize the effects of local and global earthquakes, as well as what it means for a board of chips to be finite or infinite. In Section~\ref{Stability of Finite Boards}, we introduce a potential function that maps finite board configurations to $\mathbb{R}$. We then use the fact that local earthquakes strictly increase the value of this potential function to show that finite boards cannot cycle back to previous configurations, from which we will conclude that all finite boards stabilize (reach a steady state). In Section~\ref{Acyclicity of Infinite Boards}, we modify the potential function so that it converges to a real number even for infinite boards. From this, we conclude that infinite boards with bounded chip stack size never cycle back to previous configurations. In Section~\ref{Order of Redistribution}, we show that the order in which local earthquakes happens does not matter, thus answering a question in the original problem statement.

For the remainder of the paper, we consider other approaches and extensions of the problem. In  Section~\ref{Time until Stability}, we present results from our computer simulations of repeated earthquakes. In particular, we look at the starting arrangement where all of the board's $n$ chips are initially on one central square. We conjecture based on experimental evidence that such arragnements stabilize after $\Theta(n \log n)$ global earthquakes. In Section~\ref{Board Variants: n-Trees}, we generalize away from the chessboard set up, and consider what happens when a similar process happens on tree graphs of infinite size. Finally, in Section~\ref{Computing with Chips} we present a model of computation based on the poker chip board and show how to construct simple Boolean circuits. We will also give impossibility results that show the limits of our model of computation.
\section{Definitions and Notation}
\label{Definitions and Notation}
